---
title: "Rasch_Model"
author: "Felix Eßer"
date: "8/9/2020"
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = file.path("..", "Output")) })
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The Rasch Model

The Rasch model is a one-dimensional measurement model (latent-trait model) for dichotomous items 

This document serves the purpose to illustrate the analysis of a Rasch model with marginal respect to the mathematical foundations. The focus will be on the implementation in R. 

Within the Rasch model, two model assumptions are made:

1. Rasch homogeneity:
    - All $p$ items measure the same latent variable.
    - $$ P(Y_i = 1|\eta) = \frac{e^{\eta - \alpha_i}}{1 + e^{\eta - \alpha_i}}$$
2. Local independence of items
    - The latent variable $\eta$ explains all correlations between the items.
    
    
Model assessment

1. Equality of item parameters in sub-populations
    1.1 Andersen test (Conditional likelihood-quotient-test)
    1.2 Graphical model test
    1.3 Wald-test
    1.4 Rasch-mixture-analysis
2. Global model validity
    2.1 Probability distribution of response patterns
    2.2 Likelihood-quotient-test
3. Equality of person scores in reduced Rasch models


# Packages to analyse Rasch Models

Of course there are different `R` packages available to analyse a Rasch model: 

- `eRm` - extended Rasch model
- `ltm` - latent trait model and
- `TAM` - Test Analysis Modules

There are different advantages of these packages - different analyses can be performed to assess the model validity.

```{r load packages, message=FALSE}
# Packages to perform Rasch models
library(eRm)
library(ltm)
library(TAM)

# Additional packages
library(psychomix) # to fit finite mixtures of Rasch models
library(psych) # to calculate Yules' Q
library(tidyverse) # data transformation, plotting, etc.
```


```{r mean and variance}
data(raschdat1) # load data from eRm package
raschdat1 %>% 
  summarise_all(list(Mean = mean, Variance = var)) %>% 
  pivot_longer(cols = everything(),
               names_to = c("Item", "Statistic"),
               names_pattern = "(I\\d+)_(.+)",
               values_to = "Value") %>% 
  pivot_wider(id_cols = Item,
              names_from = Statistic,
              values_from = Value)
```

```{r Yules-Q plot}
yules_mat <- matrix(nrow = ncol(raschdat1), ncol = ncol(raschdat1),
                    dimnames = list(colnames(raschdat1), colnames(raschdat1)))

for (i in 1:nrow(yules_mat)) {
  for (j in 1:ncol(yules_mat)) {
    yules_mat[i,j] <- Yule(table(raschdat1[,i], raschdat1[,j]))
  }
}

yules_mat %>% 
  # convert to tibble
  as_tibble() %>%
  # add variable names as column
  mutate(variable_1 = colnames(.)) %>%
  # transform in wide format
  pivot_longer(cols = -variable_1, 
               names_to = "variable_2", 
               values_to = "Yules_Q") %>% 
  # drop upper triangle of correlation matrix
  filter(as.integer(str_extract(variable_1, "\\d+")) <= 
           as.integer(str_extract(variable_2, "\\d+"))) %>% 
  # transform variables to factors
  mutate(variable_1 = factor(variable_1,
                             levels = str_c("I", 1:length(unique(variable_1)))),
         variable_2 = factor(variable_2,
                             levels = rev(levels(variable_1)))) %>% 
  # create plot
  ggplot(aes(x = variable_1, y = variable_2, fill = Yules_Q)) +
  geom_tile() +
  geom_text(aes(label = round(Yules_Q, 2)), size = 2) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                       midpoint = 0, limit = c(-1, 1),
                       name = "Yules' Q") +
  theme_minimal() +
  theme(axis.title = element_blank())
```


# Rasch Model with the `eRm` package

```{r Rasch model fit}
# Fit of Rasch Model with conditional maximum likelihood
rasch_model <- RM(raschdat1,
                  sum0 = TRUE) # sum0 = Normed to sum = 0; default

summary(rasch_model)
```



## Itemcharactersic function

```{r ICC}
plotjointICC(rasch_model)
#plotICC(rasch_model, ask = FALSE) # get ICC for every item
```






```{r person scores}
person_rasch <- person.parameter(rasch_model)
print(person_rasch) # max number of correct answers was 26; nobody solved no item
summary(person_rasch)
```


Person-Item-Map
```{r PI-Map}
plotPImap(rasch_model)
```



# Model assessment

Test model validity

## Equality of item parameters in sub-populations
    
### Andersen Test

```{r Andersen Test}
lr_test <- LRtest(rasch_model, splitcr = "median" ) # split criterion median is the default
summary(lr_test)
```
  
The Andersen test yielded in a non-signifikant result, $\chi^2(29) = 32.54, p = .30$. Therefore, the null-hypotheses that the itemparameters are the same can be maintained.


## Graphical model test

By means of the function `plotGOF` from the `eRm` package, it is possible to obtain a plot in which the itemparameters of the andersen test are shown in respect to the split criterion. 
    
    
```{r}
plotGOF(lr_test,
        conf = list()) # outputs confidence interval ellipses
```


## Wald-Test

The Walt-test is a test to validate for every item whether the intemparameters are the same in subpopulation (created through split criterion).
Only interpret these results if the global test (Andersen-Test) is significant. This procedure is comparable to the execution of an ANOVA. The global test has to be significant to interpret.
```{r}
wald_test <- Waldtest(rasch_model, splitcr = "median") # median is the default

# Output ordered matrix in respect to p-values
wald_test$coef.table[order(wald_test$coef.table[,"p-value"]),]
```



## Rasch-mixture-analysis

Combination of Rasch-model with latent class analysis. Sub-populations

Usage of the function `raschmix` from the package `psychomix`.

```{r rasch-mixture-analysis}
mix_1_grp <- raschmix(data = as.matrix(raschdat1), # data input has to be a matrix
                      k = 1, # number of latent classes
                      scores = "meanvar")
summary(mix_1_grp)
```

```{r}
mix_2_grp <- raschmix(data = as.matrix(raschdat1), # data input has to be a matrix
                      k = 2, # number of latent classes
                      scores = "meanvar")
summary(mix_2_grp)
```



    1.1 Graphical model test
    1.2 Conditional likelihood-quotient-test
    1.3 Wald-test
    1.4 Rasch-mixture-analysis
2. Global model validitiy
    2.1 Probability distribution of response patterns
    2.2 Likelihood-quotient-test
    
    
    
3. Equality of person scores in reduced Rasch models

```{r}
MLoef(rasch_model, splitcr = "median")
```



# Itemfit and Personfit Statistics

The `itemfit` function of the `eRm` package needs an object of class `ppar` as input. This type of object is created with the `person.parameter` function. Therefore, the object with...
```{r itemfit}
rasch_itemfit <- itemfit(person_rasch) %>% # calculates itemfit statistics
  print(visible = FALSE) # returns otherwise printed matrix

rasch_itemfit[order(rasch_itemfit[,"p-value"]),] # prints matrix ordered by p-value
```

The $\chi^2$-statistic is a residualbased 

Given model validity 5% of Person should have p-values below 5%.

The following is from the book: Applying the Rasch Model

Infit and Outfit Mean Squares are unstandardized measures of fit and Infit/Outfit t is the standardized form

Mean Squares of the fit statistic represent the mean of the squared residuals for any item. Residuals = differences between Rasch model's theoretical expectation of item performance and the observed performance for an item. 
"Mean square fit statistics indicate the size of how much misfit is revealed in the actual data"

Standardized fit statistic: mean squares are transformed regarding sample size to produce a statistic with a t distribution (acceptable values should lie between -1.96 and 1.96)

Mean squares: how much misfit; t value: how likely is the misfit


Infit statistics: information-weighted indicator of misfit; more weight is given to the performances of those persons located closer to the item's difficulty value --> because if ability of person is close to item's difficulty, performance of person gives more insight into item's performance

Outfit statistics: not weighted indicator of misfit --> more sensitive to the influence of outliers = performance of persons distant from items location

--> Therefore, often more focus on infit statistics; extreme infit statistics usually cause more concern than extreme outfit.


```{r personfit}
rasch_persfit <- personfit(person_rasch) %>% 
  print(visible = FALSE) # returns otherwise printed matrix

rasch_persfit[order(rasch_persfit[,"p-value"]),]
```



```{r person separation reliability}
sep_rel <- SepRel(person_rasch)
summary(sep_rel)
```

person separation Reliability: proportion of person variance that is not due to error. Concept of person separation reliability is very similar but not identical to reliability indices such as Cronbach's $\alpha$

In this example the person separation reliability equals `r round(sep_rel$sep.rel, 2)`



```{r}
# person separation reliability by hand
var(person_rasch$theta.table$`Person Parameter`)/(sep_rel$MSE + var(person_rasch$theta.table$`Person Parameter`))

(var(person_rasch$theta.table$`Person Parameter`) - sep_rel$MSE)/ var(person_rasch$theta.table$`Person Parameter`)

```


```{r}
# item separation reliability 

(var(rasch_model$betapar) - sep_rel$MSE)/var(rasch_model$betapar)




```






