---
title: "Rasch_Model"
author: "Felix Eßer"
date: "8/9/2020"
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = file.path("..", "Output")) })
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The Rasch Model

The Rasch model is a one-dimensional measurement model (latent-trait model) for dichotomous items 

This document serves the purpose to illustrate the analysis of a Rasch model with marginal respect to the mathematical foundations. The focus will be on the implementation in R. 

Within the Rasch model, two model assumptions are made:

1. Rasch homogeneity:
    - All $p$ items measure the same latent variable.
    - $$ P(Y_i = 1|\eta) = \frac{e^{\eta - \alpha_i}}{1 + e^{\eta - \alpha_i}}$$
2. Local independence of items
    - The latent variable $\eta$ explains all correlations between the items.
    
    
Model assessment

1. Equality of item parameters in sub-populations
    1.1 Graphical model test
    1.2 Conditional likelihood-quotient-test
    1.3 Wald-test
    1.4 Rasch-mixture-analysis
2. Global model validitiy
    2.1 Probability distribution of response patterns
    2.2 Likelihood-quotient-test
3. Equality of person scores in reduced Rasch models


# Packages to analyse Rasch Models

Of course there are different `R` packages available to analyse a Rasch model: 

- `eRm` - extended Rasch model
- `ltm` - latent trait model and
- `TAM` - Test Analysis Modules

There are different advantages of these packages - different analyses can be performed to assess the model validity.

```{r load packages, message=FALSE}
# Packages to perform Rasch models
library(eRm)
library(ltm)
library(TAM)

# Additional packages
library(psychomix) # to fit finite mixtures of Rasch models
library(psych) # to calculate Yules' Q
library(tidyverse)
```


```{r Mean and Variance}
data(raschdat1)
raschdat1 %>% 
  summarise_all(list(Mean = mean, Variance = var)) %>% 
  pivot_longer(cols = everything(),
               names_to = c("Item", "Statistic"),
               names_pattern = "(I\\d+)_(.+)",
               values_to = "Value") %>% 
  pivot_wider(id_cols = Item,
              names_from = Statistic,
              values_from = Value)
```

```{r Yules-Q}
yules_mat <- matrix(nrow = ncol(raschdat1), ncol = ncol(raschdat1),
                    dimnames = list(colnames(raschdat1), colnames(raschdat1)))

for (i in 1:nrow(yules_mat)) {
  for (j in 1:ncol(yules_mat)) {
    yules_mat[i,j] <- Yule(table(raschdat1[,i], raschdat1[,j]))
  }
}

yules_mat %>% 
  as_tibble() %>% # convert to tibble
  mutate(variable_1 = colnames(.)) %>% # add variable names as column
  pivot_longer(cols = -variable_1, # transform in wide format
               names_to = "variable_2", 
               values_to = "Yules_Q") %>% 
  filter(as.integer(str_extract(variable_1, "\\d+")) <= 
           as.integer(str_extract(variable_2, "\\d+"))) %>% # drop upper triangle of correlation matrix
  mutate(variable_1 = factor(variable_1, levels = str_c("I", 1:length(unique(variable_1)))),
         variable_2 = factor(variable_2, levels = rev(levels(variable_1)))) %>% # transform variables to factors
  ggplot(aes(x = variable_1, y = variable_2, fill = Yules_Q)) +
  geom_tile() +
  geom_text(aes(label = round(Yules_Q, 2)), size = 2) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                       midpoint = 0, limit = c(-1, 1),
                       name = "Yules' Q") +
  theme_minimal() +
  theme(axis.title = element_blank())
```


# Rasch Model with the `eRm` package

```{r Rasch model fit}
# Fit of Rasch Model with conditional maximum likelihood
rasch_model <- RM(raschdat1)

summary(rasch_model)
```

